---
title: "Initial Korean covid 19"
author: "Daniel"
date: "Fall 2023"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata <- kRdata %>%
mutate(across(where(is.character), as.factor))

```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 2020
### This data was retrieved from a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
```{r}
str(kRdata)
```

# Your Question - Is dying independent of sex?

$H_0$: There is no difference in death between male and female in rearguards of covid.

$H_A$: There is a difference in death between male and female in rearguards of covid.

#Methods

Since we are comparing $result to $sex, we then are going to use Cat~Cat helper file.

## Graphical results - use ggplot commands to do this  
```{r}
dd2 <- kRdata %>% group_by(result,Sex) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=Sex,y=count,fill=result))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=Sex,y=prcnt*100,fill=result)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

We can see that males die more often than females in this graph.

### Numerical Summaries  

```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```
  
  The data will show us that approximately, males die 2 times more often than females.

## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
```

We get a  p-value of 0.0012 which is smaller than 0.05, meaning there is a correlation between death and sex

```{r}
fisher.test(table1)
```

With this test, we get a p-value of 9e^-4 (.0009) which is smaller than 0.05 which also means a correlation between death and sex. You have a 2.3 to 1 odds of dying from covid as a male in Korea than a female.

## Conclusions

Is there a correlation between death and sex? Well with our data, we find out that there is a corrolation between death and sex regarding covid. With chisq test we get a p-value of 0.0012 which is less than 0.05 which means we reject the null hypothesis. With the fisher exact test we get odds of 2.3 times more likely to die from covid as a male in Korea.

